---
title: "PCA Scale"
author: "Jen Baron"
date: 'January 18, 2022'
output: html_document
---

#Load Packages
```{r}
library(tidyverse)
library(dplyr)
library(ggrepel)
library(truncnorm)
library(vegan)
library(ade4)
library(gridExtra) 
library(RColorBrewer)
```

# Create Data Frame with Landscape Metrics

## Historical

Variables:

- P.grassland = proportion grassland
- P.open = proportion open canopy
- P.closed = proportion closed forest
- P.dense = proportion dense forest
- N.patches = number of patches
- F.edge = amount of forest edge (m)
- P.cover = percent canopy cover (%)
- Fuel.type = fuel type
- P.size = mean patch size (m^2)
- C = contagion
- SHEI = Shannon's evenness index
- pladj = percentage of like adjacencies
- LSI = landscape shape index


Sample from normal distribution with limits (a,b), mean, and standard deviation

```{r}
set.seed(303)

P.grassland <- rtruncnorm(a=0.10, b=.50, n=91, mean = .40, sd = .10) %>% round(2)
P.open <- rtruncnorm(a=0.10, b = 0.40, n = 91, mean = 0.30, sd = 0.10) %>% round(2)
P.closed <- rtruncnorm(a=.10, b=.30, n=91, mean = .20, sd = .10) %>% round(2)
P.dense <- rtruncnorm(a=0.10, b = 0.40, n = 91, mean = 0.10, sd = 0.10) %>% round(2)

N.patches <- rtruncnorm(a=2, b=15, n=91, mean = 3, sd = 5) %>% round(0)
F.edge <- rtruncnorm(a=100, b=1000, n=91, mean = 500, sd = 300) %>% round(2)
P.cover <- rtruncnorm(a=10, b=90, n=91, mean = 30, sd = 50) %>% round(2)
#Fuel.type <- c("C3", "O1-a", "C4", "O1-b", "C7", "O1-a", "C5", "C6", "C7", "C5")
#P.size <- rtruncnorm(a=60, b = 450, n = 91, mean = 300, sd = 180) %>% round (0) 
C <- rtruncnorm(a=0, b = 1, n = 91, mean = 0.5, sd = 0.3) %>% round (2) 
SHEI <- rtruncnorm(a=0, b = 1, n = 91, mean = 0.5, sd = 0.15)
#pladj <- rtruncnorm(a=0, b = 1, n = 91, mean = 0.5, sd = 0.15)
LSI <- rtruncnorm(a=1, b = 10, n = 91, mean = 5, sd = 3)
```


```{r}
structure.H <-
  data.frame(
  "site" = 1:91,
  "p.grassland" = P.grassland,
  "p.open" = P.open,
  "p.closed" = P.closed,
  "p.dense" = P.dense,
  "n.patches" = N.patches,
  "f.edge" = F.edge,
  "p.cover" = P.cover, 
  "C" = C,
  "SHEI" = SHEI,
  "LSI" = LSI)

structure.H$time <- "Historical"

head(structure.H)
```
 
## Modern

```{r}
set.seed(130)

structure.M <- structure.H

grass.change <- (sample(c(20:50), size = 91, replace = TRUE))*0.01
structure.M$p.grassland <- structure.H$p.grassland - grass.change    
structure.M$p.grassland[structure.M$p.grassland<0] <- 0

open.change <- (sample(c(5:20), size = 91, replace = TRUE))*0.01
structure.M$p.open <- structure.H$p.open - open.change 
structure.M$p.open[structure.M$p.open<0] <- 0

closed.change <- (sample(c(5:20), size = 91, replace = TRUE))*0.01
structure.M$p.closed <- structure.H$p.closed + closed.change    

dense.change <- (sample(c(15:30), size = 91, replace = TRUE))*0.01
structure.M$p.dense <- structure.H$p.dense + dense.change    

patches.change <- (sample(c(-5:15), size = 91, replace = TRUE)) 
structure.M$n.patches <- structure.H$n.patches + patches.change  

edge.change <- (sample(c(-100:300), size = 91, replace = TRUE)) 
structure.M$f.edge <- structure.H$f.edge + edge.change

cover.change <- (sample(c(-10:50), size = 91, replace = TRUE)) 
structure.M$p.cover <- structure.H$p.cover + cover.change
structure.M$p.cover[structure.M$p.cover>95] <- 95

# size.change <- (sample(c(20:300), size = 91, replace = TRUE))
# structure.M$p.size <- structure.H$p.size + size.change

C.change <- (sample(c(0:30), size = 91, replace = TRUE))*0.01
structure.M$C <- structure.H$C + C.change
  
SHEI.change <- (sample(c(20:50), size = 91, replace = TRUE))*0.01
structure.M$SHEI <- structure.H$SHEI + SHEI.change

# pladj.change <- (sample(c(20:45), size = 91, replace = TRUE))*0.01
# structure.M$pladj <- structure.H$pladj + pladj.change

LSI.change <-(sample(c(0:7), size = 91, replace = TRUE))
structure.M$LSI <- structure.H$LSI - LSI.change

structure.M$time <- "Modern"

head(structure.M)
```

## Join Historcal & Modern

```{r}
structure <- bind_rows(structure.H, structure.M)

structure$time <- as.factor(structure$time)

str(structure)
head(structure)
```

# Run PCA

```{r}
structure %>% select(p.grassland,
                     p.open,
                     p.closed,
                     p.dense,
                     n.patches,
                     f.edge,
                     p.cover,
                     C,
                     SHEI,
                     LSI) -> PCA.data

prcomp(PCA.data, scale. = TRUE) -> pca.1

summary(pca.1)

biplot(pca.1)
```

```{r}
PC1 <- predict(pca.1)[,1]
PC2 <- predict(pca.1)[,2]

rescale.U <- rep(pca.1$sdev, each = length(pca.1$sdev)) #get lengths

U.scale2 <- pca.1$rotation * rescale.U #multiply lengths by sqrt SD

round(U.scale2^2,2) #variability in each variable for each PC
```

```{r}
U <- data.frame(pca.1$rotation)
colnames(U) <- colnames(pca.1$rotation)
rownames(U) <- rownames(pca.1$rotation)
U$descriptors <- rownames(U)
F.1 <- data.frame(pca.1$x) 
colnames(F.1) <- colnames(pca.1$x)
rownames(F.1) <- rownames(pca.1$x)
str(U)
```

```{r}
str(structure)
F.1$site <- structure$site
F.1$time <- structure$time
F.1$fuel.type <- structure$fuel.type

str(F.1)
```

## Bi-Plots


```{r}
U$descriptors <- as.factor(U$descriptors)
levels(U$descriptors) 
levels(U$descriptors) <- c("Contagion", "Forest edge", 
                           "Landscape shape index", 
                           "Number of patches",
                           "Proportion closed forest (C-3)",
                           "Percent canopy cover",
                           "Proportion dense forest (C-4)",
                           "Proportion grassland (O1-b)",
                           "Proportion open forest (C-7)",
                           "Shannon's evenness index")
                           
biplot1 <- ggplot(F.1, aes(x = PC1, y =PC2)) + 
  geom_point(aes(col = time), size = 1, alpha = 0.3) +
  stat_ellipse(aes(x=PC1, y = PC2, col=time)) +
  theme_classic() +
  coord_fixed() +
  labs(x = 'Principle Component 1', y = "Principle Component 2") +
    scale_colour_manual(values = c("darkgreen", "darkgoldenrod4")) +
  geom_segment(data = U, aes(xend = PC1*5, yend = PC2*5,x = 0, y = 0), col = "black", 
               alpha = 0.8, arrow =arrow(length = unit(0.4, "cm"))) +
  geom_label_repel(data = U, aes(x = PC1*5, y = PC2*5, label = descriptors),
                   col = "black", nudge_y = -0.55, 
                   segment.colour = NA, size = 2.5, alpha = 0.8) +
  annotate("text", label = "Modern (2016)",
             x=2, y=2.9, size = 3.5) +
    annotate("text", label = "Historical (1950)",
             x=-2.1, y=2.5, size=3.5) +
  scale_x_continuous(breaks = c(-4, -2, 0, 2, 4)) +
  theme(legend.position="none", legend.box = "horizontal", plot.margin=grid::unit(c(0,0,0,0), "mm")) +
  guides(fill=guide_legend(title=NULL))
biplot1
```



```{r}
ggsave('figures/2_PCA_biplot_91.jpeg',biplot1, units = 'cm', width = 15, height = 11)
```


#Save Outputs

```{r}
write.csv(structure, "outputs/structure_91.csv")
```


# Reproducibility

```{r}
git2r::repository()
Sys.time()
sessionInfo()
```