---
title: "Chapter 4"
author: "Jen Baron"
date: "1/20/2022"
output: html_document
---

#Load Packages

```{r}
library(ggplot2)
library(dplyr)
library(DiagrammeR)
library(DiagrammeRsvg)
library(rsvg)
library(ggprism)
library(extrafont)
#loadfonts(); windowsFonts()
```
# Conceptual model

Add:

- Fire suppression
- Additional outputs
- Specific scenarios
- Cumulative fire risk assessment

D [label = 'Fire Suppression']
f [label = 'Flame length probability']
g [label = 'Fire line intensity']

b2c [label = 'Increase\n lightning ignitions']
b2d [label = 'Decrease\n human ignitions']

b3c [label = '2050 RCP2.6']
b3d [label = '2050 RCP8.5']

## Simple

```{r}
simple <- grViz("
digraph boxes_and_circles {
  # a 'graph' statement
  graph [overlap = true, fontsize = 12 compound=true] #compound=true to add arrows from clusters

  # several 'node' statements

node [shape = plaintext, fontname = Georgia, fontsize=18]
A [label = 'Modern']
B [label = 'Historical']

node [shape = circle, fixedsize = true, width = 1.7, fontsize=14]
a1 [label = 'Modern\n fuels']
a2 [label = 'Modern\n ignitions']
a3 [label = 'Modern\n weather']

b1 [label = 'Historical\n fuels']
b2 [label = 'Historical\n ignitions']
b3 [label = 'Historical\n weather']

node [shape = circle, fixedsize = true, width = 2, fontsize=14]
C [label = 'Wildfire\n simulation\n modelling']

node [shape = box, fixedsize = true, width = 2]
d [label = 'Fire suppression']

node [shape = box, fixedsize = true, width = 1]
d1 [label = '0%']
d2 [label = '50%']
d3 [label = '90%']

# subgraphs

subgraph cluster_0 {
		style=filled;
		color=lightgrey;
		node [style=filled,color=white];
		a1 a2 a3;}

subgraph cluster_1 {
		style=filled;
		color=lightgrey;
		node [style=filled,color=white];
		b1 b2 b3;}

subgraph cluster_d {
		style=filled;
		color=lightgrey;
		node [style=filled,color=white];
    d d1 d2 d3;}

# several 'edge' statements

A-> a2 [lhead=cluster_0]
B-> b2 [lhead=cluster_1]
a1->d [ltail=cluster_0]
b3->d [ltail=cluster_1]
d->d1 d->d2 d->d3
d2->C [ltail=cluster_d]

}
")
simple
```

```{r}
simple %>%
    export_svg() %>%
    charToRaw %>%
    rsvg_png("figures/5_simulation_simple2.png")
```

## Restoration to historical fuels

```{r}
grViz("
digraph boxes_and_circles {
  # a 'graph' statement
  graph [overlap = true, fontsize = 12 compound=true] #compound=true to add arrows from clusters

  # several 'node' statements
node [shape = circle, fixedsize = true, width = 1.7]
a1 [label = 'Modern\n fuels']
a2 [label = 'Modern\n ignitions']
a3 [label = 'Modern\n weather']

b1 [label = 'Historical\n fuels']
b2 [label = 'Historical\n ignitions']
b3 [label = 'Historical\n weather']

c1 [label = 'Historical\n fuels']
c2 [label = 'Modern\n ignitions']
c3 [label = 'Modern\n weather']

node [shape = circle, fixedsize = true, width = 2]
G [label = 'Wildfire\n simulation\n modelling']

node [shape = box, fixedsize = true, width = 2]
d [label = 'Fire suppression']

node [shape = box, fixedsize = true, width = 1]
d1 [label = '0%']
d2 [label = '50%']
d3 [label = '90%']

# subgraphs

subgraph cluster_0 {
		style=filled;
		color=lightgrey;
		node [style=filled,color=white];
		a1 a2 a3;}

subgraph cluster_1 {
		style=filled;
		color=lightgrey;
		node [style=filled,color=white];
		b1 b2 b3;}

subgraph cluster_2 {
		style=filled;
		color=lightgrey;
		node [style=filled,color=white];
		c1 c2 c3;}

subgraph cluster_d {
		style=filled;
		color=lightgrey;
		node [style=filled,color=white];
    d d1 d2 d3;}

# several 'edge' statements

a1->d [ltail=cluster_0]
b2->d [ltail=cluster_1]
c3->d [ltail=cluster_2]
d->d1 d->d2 d->d3
d2->G [ltail=cluster_d]


}
")
```


## Complex

```{r}
grViz("
digraph boxes_and_circles {
  # a 'graph' statement
  graph [overlap = true, fontsize = 12 compound=true] #compound=true to add arrows from clusters

  # several 'node' statements

node [shape = plaintext, fontname = Georgia]

node [shape = circle, fixedsize = true, width = 1.7]
b1 [label = 'Fuels']
b2 [label = 'Ignitions']
b3 [label = 'Weather']
b4 [label = 'Topography']

#Alternate scenarios
node [shape = circle, fixedsize = true, width = 1]

b1a [label = 'Historical\n (1950)']
b1b [label = 'Modern\n (2021)']
b1c [label = 'Future']
b1d [label = 'Thinning\n & tending']
b1e [label = 'Harvesting']
b1f [label = 'Prescribed\n fire']
b1g [label = 'Ecological\n restoration']

b2a [label = 'Historical\n (1950)']
b2b [label = 'Modern\n (2021)']


b3a [label = 'Historical\n (1950)']
b3b [label = 'Modern\n (2021)']


node [shape = circle, fixedsize = true, width = 2]
C [label = 'Wildfire\n simulation\n modelling']


node [shape = box, fixedsize = true, width = 2]
d [label = 'Fire suppression']

node [shape = box, fixedsize = true, width = 1]
d1 [label = '0%']
d2 [label = '50%']
d3 [label = '90%']

# subgraphs


subgraph cluster_1 {
		style=filled;
		color=lightgrey;
		node [style=filled,color=white];
		b4 b1 b2 b3;}

subgraph cluster_b1 {
		style=filled;
		color=lightgrey;
		node [style=filled,color=white];
		b1a b1b b1c b1d b1e b1f b1g;}

subgraph cluster_b2 {
		style=filled;
		color=lightgrey;
		node [style=filled,color=white];
		b2a b2b;}

subgraph cluster_b3 {
		style=filled;
		color=lightgrey;
		node [style=filled,color=white];
		b3a b3b;}

subgraph cluster_d {
		style=filled;
		color=lightgrey;
		node [style=filled,color=white];
    d d1 d2 d3;}


# several 'edge' statements


b2b->b2 [ltail=cluster_b2]
b3b->b3 [ltail=cluster_b3]
b1b-> b1[ltail=cluster_b1]
b1c->b1d b1c->b1e b1c->b1f b1c->b1g


b2->d [ltail=cluster_1]
d2->C [ltail=cluster_d]
d->d1 d->d2 d->d3

}
")
```
#
```{r}
grViz("
digraph boxes_and_circles {
  # a 'graph' statement
  graph [overlap = true, fontsize = 12 compound=true] #compound=true to add arrows from clusters

  # several 'node' statements

node [shape = plaintext, fontname = Georgia]

node [shape = circle, fixedsize = true, width = 1.7]
b1 [label = 'Fuels']
b2 [label = 'Ignitions']
b3 [label = 'Weather']
b4 [label = 'Topography']

#Alternate scenarios
node [shape = circle, fixedsize = true, width = 1]

b1a [label = 'Historical\n (1950)']
b1b [label = 'Modern\n (2021)']
b1c [label = 'Future']
b1d [label = 'Thinning\n & tending']
b1e [label = 'Harvesting']
b1f [label = 'Prescribed\n fire']
b1g [label = 'Ecological\n restoration']

b2a [label = 'Historical\n (1950)']
b2b [label = 'Modern\n (2021)']


b3a [label = 'Historical\n (1950)']
b3b [label = 'Modern\n (2021)']


node [shape = circle, fixedsize = true, width = 2]
C [label = 'Wildfire\n simulation\n modelling']


node [shape = box, fixedsize = true, width = 2]
d [label = 'Fire suppression']

node [shape = box, fixedsize = true, width = 1]
d1 [label = '0%']
d2 [label = '50%']
d3 [label = '90%']

# subgraphs


subgraph cluster_b1 {
		style=filled;
		color=lightgrey;
		node [style=filled,color=white];
		b1 b1a b1b b1c b1d b1e b1f b1g;}

subgraph cluster_b2 {
		style=filled;
		color=lightgrey;
		node [style=filled,color=white];
		b2 b2a b2b;}

subgraph cluster_b3 {
		style=filled;
		color=lightgrey;
		node [style=filled,color=white];
		b3 b3a b3b;}

subgraph cluster_b4 {
		style=filled;
		color=lightgrey;
		node [style=filled,color=white];
		b4;}

subgraph cluster_d {
		style=filled;
		color=lightgrey;
		node [style=filled,color=white];
    d d3 d2 d1;}

# several 'edge' statements

b1->b1a b1->b1b b1->b1c
b1c->b1d b1c->b1e b1c->b1f b1c->b1g
b1->d [ltail=cluster_b1]
 
b2->b2a b2->b2b
b2->d [ltail=cluster_b2]

b3->b3a b3->b3b
b3->d [ltail=cluster_b3]
 
b4->d [ltail=cluster_b4]

d->d1 d->d2 d->d3
d3->C [ltail=cluster_d]

}
")
```
## Parisien et al. 2019

```{r}
grViz("
digraph boxes_and_circles {
  # a 'graph' statement
  graph [overlap = true, fontsize = 12 compound=true] #compound=true to add arrows from clusters

  # several 'node' statements

node [shape = plaintext,
        fontname = Georgia]
A [label = 'Baseline conditions assessment']
B [label = 'Alternate state assessment']

node [shape = circle, fixedsize = true, width = 1.7]
a1 [label = 'Fuels & topography']
a2 [label = 'Ignitions']
a3 [label = 'Weather']

b1 [label = 'Alternate\n fuels & topography']
b2 [label = 'Alternate\n Ignitions']
b3 [label = 'Alternate\n Weather']

node [shape = circle, fixedsize = true, width = 2]
C [label = 'Wildfire\n simulation\n modelling']

node [shape = circle, fixedsize = true, width = 1.7]
c1 [label = 'Fire perimeters']
c2 [label = 'Ignitions']
c3 [label = 'Burn probability']
c4 [label = 'Fire intensity']

# subgraphs

subgraph cluster_0 {
		style=filled;
		color=lightgrey;
		node [style=filled,color=white];
		a1 a2 a3;}

subgraph cluster_1 {
		style=filled;
		color=lightgrey;
		node [style=filled,color=white];
		b1 b2 b3;}

subgraph cluster_2 {
		style=filled;
		color=lightgrey;
		node [style=filled,color=white];
		c1 c2 c3 c4;}

# several 'edge' statements

A->a2 [lhead=cluster_0]
B->b2 [lhead=cluster_1]
a1->C [ltail=cluster_0]
b3->C [ltail=cluster_1]
C->c2 [lhead=cluster_2]
}
")
```


# Reproducibility

```{r}
git2r::repository()
Sys.time()
sessionInfo()
```
