---
title: "Mini-Proposal 2"
author: "Jen Baron"
date: 'October 17, 2019'
output: html_document
---

#Load Packages
```{r}
library(tidyverse)
library(dplyr)
library(ggrepel)
library(truncnorm)
library(vegan)
library(ade4)
library(gridExtra) 
library(RColorBrewer)
```

# Create Data Frame with Landscape Metrics

## Historical

Variables:

- P.grassland = proportion grassland
- P.closed = proportion closed canopy
- P.open = proportion open canopy
- N.patches = number of patches
- F.edge = amount of forest edge (m)
- T.density = tree density (basal area / ha?)
- P.cover = percent canopy cover (%)
- Fuel.type = fuel type
- Snags = estimated number of snags ( / ha)
- T.type = dominant tree type (B = broadleaf, N = needeleaf, M = mixed)
- P.size = mean patch size (m^2)
- C = contagion
- FSS = fire sensitive species
- FAS = fire adapted species



Sample from normal distribution with limits (a,b), mean, and standard deviation

```{r}
set.seed(303)
P.grassland <- rtruncnorm(a=0, b=.60, n=1000, mean = .10, sd = .20) %>% round(2)
P.closed <- rtruncnorm(a=.10, b=.60, n=1000, mean = .40, sd = .20) %>% round(2)
P.open <- rtruncnorm(a=0, b = 0.6, n = 1000, mean = 0.3, sd = 0.20) %>% round(2)
N.patches <- rtruncnorm(a=2, b=15, n=1000, mean = 3, sd = 5) %>% round(0)
F.edge <- rtruncnorm(a=100, b=1000, n=1000, mean = 500, sd = 300) %>% round(2)
T.density <- rtruncnorm(a=10, b=100, n=1000, mean = 30, sd = 40) %>% round(2)
P.cover <- rtruncnorm(a=10, b=90, n=1000, mean = 30, sd = 50) %>% round(2)
#Fuel.type <- c("C3", "O1-a", "C4", "O1-b", "C7", "O1-a", "C5", "C6", "C7", "C5")
#Snags <- rtruncnorm(a=1, b=30, n=1000, mean = 12, sd = 10) %>% round(2)
#T.type <- sample(c("B", "N", "N", "N", "M"), size = 10, replace = TRUE)
P.size <- rtruncnorm(a=60, b = 450, n = 1000, mean = 300, sd = 180) %>% round (0) 
C <- rtruncnorm(a=0, b = 1, n = 1000, mean = 0.5, sd = 0.3) %>% round (2) 
FSS <- rtruncnorm(a = 0, b = 1, n = 1000, mean = 0.4, sd = 0.2) %>% round(2)
FAS <- rtruncnorm(a = 0, b = 1, n = 1000, mean = 0.6, sd = 0.2) %>% round(2)
```


```{r}
structure.H <-
  data.frame(
  "site" = 1:1000,
  "p.grassland" = P.grassland,
  "p.closed" = P.closed,
  "p.open" = P.open,
  "p.size" = P.size,
  "n.patches" = N.patches,
  "C" = C,
  "FAS" = FAS,
  "FSS" = FSS,
  "f.edge" = F.edge,
  "p.cover" = P.cover,
  "t.density" = T.density
  )

structure.H$time <- "Historical"

head(structure.H)
```
 
## Modern

```{r}
set.seed(130)

structure.M <- structure.H

grass.change <- (sample(c(0:50), size = 1000, replace = TRUE))*0.01
structure.M$p.grassland <- structure.H$p.grassland - grass.change    
structure.M$p.grassland[structure.M$p.grassland<0] <- 0

closed.change <- (sample(c(0:27), size = 1000, replace = TRUE))*0.01
structure.M$p.closed <- structure.H$p.closed + closed.change    

open.change <- (sample(c(0:20), size = 1000, replace = TRUE))*0.01
structure.M$p.open <- structure.H$p.open - open.change    

size.change <- (sample(c(0:300), size = 1000, replace = TRUE))
structure.M$p.size <- structure.H$p.size + size.change

patches.change <- (sample(c(-5:15), size = 1000, replace = TRUE)) 
structure.M$n.patches <- structure.H$n.patches + patches.change  

C.change <- (sample(c(0:30), size = 1000, replace = TRUE))*0.01
structure.M$C <- structure.H$C + C.change
  
FAS.change <- (sample(c(0:30), size = 1000, replace = TRUE))*0.01
structure.M$FAS <- structure.H$FAS - FAS.change
  
FSS.change <-(sample(c(0:25), size = 1000, replace = TRUE))*0.01
structure.M$FSS <- structure.H$FSS + FSS.change

edge.change <- (sample(c(-100:300), size = 1000, replace = TRUE)) 
structure.M$f.edge <- structure.H$f.edge + edge.change

cover.change <- (sample(c(-10:50), size = 1000, replace = TRUE)) 
structure.M$p.cover <- structure.H$p.cover + cover.change
structure.M$p.cover[structure.M$p.cover>95] <- 95

density.change <- (sample(c(-10:50), size = 1000, replace = TRUE)) 
structure.M$t.density <- structure.H$t.density + density.change


#structure.M$fuel.type <- c("C3", "C6", "C4", "C4", "C7", "O1-a", "C4", "C6", "C7", "C6")

#snags.change <- (sample(c(-5:30), size = 1000, replace = TRUE)) 
#structure.M$snags <- structure.H$snags - snags.change 
#structure.M$snags[structure.M$snags<0] <- 0

#structure.M$t.type <- sample(c("B", "N", "N", "N", "N", "N", "N", "M"), size = 10, replace = TRUE)


structure.M$time <- "Modern"

head(structure.M)
```

## Join Historcal & Modern

```{r}
structure <- bind_rows(structure.H, structure.M)

#structure$fuel.type <- as.factor(structure$fuel.type)
#structure$t.type <- as.factor(structure$t.type)
structure$time <- as.factor(structure$time)

str(structure)
head(structure)
```

# Run PCA

```{r}
structure %>% select(p.grassland, p.closed, p.open, p.size, n.patches, C, FAS, FSS, f.edge, p.cover, t.density) -> PCA.data

prcomp(PCA.data, scale. = TRUE) -> pca.1

summary(pca.1)

biplot(pca.1)
```

```{r}
PC1 <- predict(pca.1)[,1]
PC2 <- predict(pca.1)[,2]

rescale.U <- rep(pca.1$sdev, each = length(pca.1$sdev)) #get lengths

U.scale2 <- pca.1$rotation * rescale.U #multiply lengths by sqrt SD

round(U.scale2^2,2) #variability in each variable for each PC
```

```{r}
U <- data.frame(pca.1$rotation)
colnames(U) <- colnames(pca.1$rotation)
rownames(U) <- rownames(pca.1$rotation)
U$descriptors <- rownames(U)
F.1 <- data.frame(pca.1$x) 
colnames(F.1) <- colnames(pca.1$x)
rownames(F.1) <- rownames(pca.1$x)
str(U)
```

```{r}
str(structure)
F.1$site <- structure$site
F.1$time <- structure$time
F.1$fuel.type <- structure$fuel.type

str(F.1)
```

## Bi-Plots

```{r}
U$descriptors <- as.factor(U$descriptors)
levels(U$descriptors) 
levels(U$descriptors) <- c("Contagion", "Forest Edge", "Fire-Adapted Species", "Fire Sensitive Species",
                           "Number of Patches", "Proportion Closed Canopy", 
                           "Percent Canopy Cover", "Proportion Grassland", "Proportion Open Forest",
                           "Mean Patch Size", "Tree Density")

biplot1 <- ggplot(F.1, aes(x = PC1, y =PC2)) + 
  geom_point(aes(col = time), size = 1, alpha = 0.3) +
  stat_ellipse(aes(x=PC1, y = PC2, col=time)) +
  theme_classic() +
  coord_fixed() +
  labs(x = 'Principle Component 1', y = "Principle Component 2") +
    scale_colour_manual(values = c("darkgreen", "darkgoldenrod4")) +
  geom_segment(data = U, aes(xend = PC1*5, yend = PC2*5,x = 0, y = 0), col = "black", 
               alpha = 0.8, arrow =arrow(length = unit(0.4, "cm"))) +
  geom_label_repel(data = U, aes(x = PC1*5, y = PC2*5, label = descriptors),
                   col = "black", nudge_y = -0.4, 
                   segment.colour = NA, size = 2.5, alpha = 0.8) +
  scale_x_continuous(breaks = c(-4, -2, 0, 2, 4)) +
  theme(legend.position="none", legend.box = "horizontal", plot.margin=grid::unit(c(0,0,0,0), "mm")) +
  guides(fill=guide_legend(title=NULL))
biplot1
```



```{r}
ggsave('fig.1.jpeg',biplot1, units = 'cm', width = 15, height = 11)
```


# Create Landscape Change Index 

```{r}
landscape <- F.1 %>% select(PC1, PC2, site, time)

landscape.H <-landscape %>% filter(time == "Historical")

landscape.M <-landscape %>% filter(time == "Modern")

LCI.1 <- landscape.M$PC1 - landscape.H$PC1
LCI.2 <- landscape.M$PC2 - landscape.H$PC2
site <- landscape.H$site

LC <- data.frame(site, LCI.1, LCI.2) %>% round(4)
```

Think about how to turn these values into one value in polar co-ordinates 

## Create Matrix with Constraining Variables

```{r}
set.seed(210)
Elevation <- rtruncnorm(a=100, b=4000, n=1000, mean = 2000, sd = 1000) %>% round(2)
Summer.temp <- rtruncnorm(a=14, b=20, n=1000, mean = 18, sd = 6) %>% round(2)
Winter.temp <-rtruncnorm(a=-10, b=10, n=1000, mean = -3, sd = 6) %>% round(2)
Summer.precip <- rtruncnorm(a=300, b=600, n=1000, mean = 500, sd = 200) %>% round(2)
Winter.precip <- rtruncnorm(a=100, b=400, n=1000, mean = 200, sd = 200) %>% round(2)
Warming <- rtruncnorm(a=-1, b=3, n=1000, mean = 1.9, sd = 0.5) %>% round(2)
Aspect <- sample(c(0:360), size = 1000, replace = TRUE)
Slope <- rtruncnorm(a=0, b=80, n=1000, mean = 50, sd = 80) %>% round(1)
#Soil <-
#Outflows <- rtruncnorm(a=0, b=80, n=10, mean = 50, sd = 80) %>% round(1)
Urban <-  rtruncnorm(a=1, b=80, n=1000, mean = 20, sd = 30) %>% round(2)
Population <- rtruncnorm(a=0, b=1300, n=1000, mean = 5, sd = 100) %>% round(0)
Fire <- rtruncnorm(a=10, b=200, n=1000, mean = 70, sd = 80) %>% round(2)
Transport <-rtruncnorm(a=0, b=14, n=1000, mean = 1.2, sd = 3) %>% round(2)
HFI <- rtruncnorm(a=0, b=60, n=1000, mean = 30, sd = 20) %>% round(0)
Cutblocks <- rtruncnorm(a=0, b=300, n=1000, mean = 55, sd = 70) %>% round(1)
```

```{r}
env <-
  data.frame(
  "site" = 1:1000,  
#topographic
  "elevation" = Elevation,
  "aspect" = Aspect,
  "slope" = Slope,
#climatic
  "summer.temp" = Summer.temp,
  "winter.temp" = Winter.temp,
  "summer.precip" = Summer.precip,
  "winter.precip" = Winter.precip,
  "warming" = Warming,
#anthropogenic
  "population" = Population,
  "urban" = Urban,
  "fire" = Fire,
  "transport" = Transport,
  "HFI" = HFI,
  "cutblocks" = Cutblocks
  )
head(env)
```

```{r}
#write.csv(structure, "structure.csv")
#write.csv(env, "environment.csv")
```

```{r}
head(structure)
head(environment)
```


```{r}
structure.summary <- structure %>% group_by(time) %>%
  summarize_at(vars(p.grassland:t.density), mean)
structure.summary

historical.mean <- structure.summary %>% filter(time == "Historical") %>% select(-time)
historical.mean <- gather(historical.mean, "metric") 
names(historical.mean)[names(historical.mean)=="value"] <- "Historical"

modern.mean <- structure.summary %>% filter(time == "Modern") %>% select(-time)
modern.mean <- gather(modern.mean, "metric") 
names(modern.mean)[names(modern.mean)=="value"] <- "Modern"

summary.mean <- left_join(historical.mean, modern.mean)
summary.mean

#write.csv(summary.mean, "summarize.csv")
```

```{r}
head(env)
env.summary <- env %>% summarize_at(vars(elevation:cutblocks), mean)
env.summary <- gather(env.summary)
env.summary

#write.csv(env.summary, "env.summary.csv")
```




```{r}
LC <- left_join(LC, env, by = "site")
```



```{r}
str(LC)
model.1 <- lm(data = LC, LCI.1 ~ elevation + warming + aspect + slope)
summary(model.1)
```















#RDA

```{r}
head(landscape)
```


## Create Matrix of Change in Landscape

```{r}
landscape <- structure.M

landscape$D.grass <- structure.H$p.grassland - structure.M$p.grassland
landscape$D.closed <- structure.H$p.closed - structure.M$p.closed
landscape$D.open <- structure.H$p.open - structure.M$p.open
landscape$D.size <- structure.H$p.size - structure.M$p.size
landscape$D.patches <- structure.H$n.patches - structure.M$n.patches
landscape$D.C <- structure.H$C - structure.M$C
landscape$D.FAS <- structure.H$FAS - structure.M$FAS
landscape$D.FSS <- structure.H$FSS - structure.M$FSS
landscapeD.edge <- structure.H$f.edge - structure.M$f.edge
landscape$D.cover <- structure.H$p.cover - structure.M$p.cover
landscape$D.density <- structure.H$t.density - structure.M$t.density

landscape <- landscape %>% select(D.grass, D.closed, D.open, D.size, D.patches, D.C, D.FAS, D.FSS, D.edge, D.density, D.cover, -site)

head(landscape)
```

```{r}
env <- env %>% select(-site)
head(env)
```


The optional argument scale. = TRUE standardized the variables before conducting PCA. This is important if the units differ among variables, because we do not want the variables with the larger numbers to dominate the analysis. I don't know how to scale in capscale so I'm going to pre-standardize the variables using scale. 

```{r}
scale(landscape, center = TRUE, scale = TRUE) %>% round(2) -> landscape
rda.1 <- capscale(landscape ~ elevation + aspect + slope + summer.temp + winter.temp + summer.precip + winter.precip + warming + population + urban + fire + transport + HFI + cutblocks, data = env) #must use distance = ""
```

```{r}
plot(rda.1)
```

**Model Selection**

Using backwards selection 

```{r}
set.seed(300)
rda.2 <- ordistep(rda.1)
```

**Check for colinearity**

```{r}
vif.cca(rda.2)
```

Everything ~1, nothing > 10, looks good.


**Check Axes**

```{r}
set.seed(212)
anova(rda.2, by = "axis")
```

**Final Model**

```{r}
set.seed(212)
ordistep(rda.2)
```

```{r}
rda.2 <- rda.1
```

```{r}
rda.scores <- scores(rda.2, display = c("wa", "bp", "sp")) #extract scores

rda.sp <- as.data.frame(rda.scores$species)
rda.sp$names <- rownames(rda.sp)

rda.site <- as.data.frame(rda.scores$sites) #site scores
rda.site$names <- rownames(rda.site)
rda.site$elevation <- env$elevation 
rda.site$warming <- env$warming
str(rda.site)

rda.env <- as.data.frame(rda.scores$biplot) 
rda.env$names <- rownames(rda.env)

rda.env$names
rda.sp$names <- as.factor(rda.sp$names)
levels(rda.sp$names) <- c("Contagion", "Forest Edge", "Fire-Adapted Species", "Fire Sensitive Species",
                           "Number of Patches", "Proportion Closed Canopy", 
                           "Percent Canopy Cover", "Proportion Grassland", "Proportion Open Forest",
                           "Mean Patch Size", "Tree Density" )

rda.env$names <- as.factor(rda.env$names)
rda.env.T <- rda.env %>% filter(names == "elevation" | names == "aspect" | names == "slope")
rda.env.C <- rda.env %>% filter(names == "summer.temp" | names == "summer.precip"| names == "winter.temp" | names == "winter.precip" | names == "warming")
rda.env.A <- rda.env %>% filter(names == "cutblocks" | names == "fire" | names == "HFI" | names == "population" | names == "transport"| names == "urban")

rda.env.T$factor <- "Topographic"
rda.env.C$factor <- "Climatic"
rda.env.A$factor <- "Anthropogenic"
rda.env <- bind_rows(rda.env.T, rda.env.C, rda.env.A)
levels(rda.env$names) <- c("Aspect", "Distance to Cutblocks", "Elevation", "Fire Suppression", 
                           "Human Footprint Index", "Population", "Slope", "Precipitation (Summer)", 
                           "Temperature (Summer)", "Transport Density", "Distance to Urban", "Warming", 
                           "Precpitation (Winter)", "Temperature (Summer)")

```

```{r}

fig.2 <- ggplot() +
  geom_point(data = rda.site, aes(x = CAP1, y = CAP2), size = 1.5, alpha = 0.3) +
  #geom_label_repel(data = rda.sp, aes(x = CAP1, y = CAP2, label = names), size = 2, col = "springgreen4", alpha = 0.8) +
 geom_segment(data = rda.env, aes(x = 0, xend =  9*CAP1, y = 0, yend = 9*CAP2), col = "black", alpha = 0.9, size = 0.6, arrow = arrow(length = unit(0.3, "cm"))) +
  geom_label_repel(data = rda.env, aes(x = 9*CAP1 , y = 9*CAP2, label = names, fill = factor),  size = 3.5, alpha = 0.9) +
  scale_fill_brewer(palette = "Dark2") +
  labs(x = "Axis 1", y = "Axis 2") +
  theme_classic() + 
  scale_y_continuous(breaks = c(-6, -3, 0, 3, 6)) +
    scale_x_continuous(breaks = c(-4, -2, 0, 2, 4)) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"), legend.position = "none")
fig.2

```

```{r}
#ggsave('fig.2.jpeg',fig.2, units = 'cm', width = 17, height = 13)
```

